`cesR`: A package to make using the Canadian Election Study a little easier

### 2020-08-27 paper.Rmd
updated references

An election season brings with it a stream of predictions made using a variety of resources and methods. While election predictions can provide insight as to the direction of an election and voter intention during an election, a retrospective on an election can be equally enlightening. Survey data of this sort can provide a better understanding of an election after-the-fact and can provide clarity as to the changes in the political climate over the years. The Canadian Election Study (CES) is a series of large-scale surveys conducted every election year that seeks to provide such a retrospective understanding to Canadian elections. By providing data on Canadian elections since 1965 up to, most recently, 2019 the CES provides insight into such topics as the intentions of voters, what issues voters deem important, and the perception of parties and candidates (Canadian Election Study, 2019). For example, @bittner2018leaders used CES data to assess the level of presidentialization experienced in Canadian elections over time; while other studies have used CES data to assess the level of Islamophobia in Canada [@wilkins2018islamophobia], how anti-party rhetoric affects voter behaviour [@gidengil2001correlates], and how political trust influences voter behaviour [@BelangerNadeau2005]. The collection of and access to such data is integral to this research. The R package `cesR` ('caesar'), as presented in this paper, seeks to complement the work of the CES, and enhance social science research and the teaching of statistical processes using survey data by providing a means for R users to easily access CES survey datasets.

The CES survey datasets are publicly available from various repositories across the internet in various data types (Canadian Election Study, 2019; UBC, n.d.; Odesi, n.d.). As the ability to find and access data is integral to quality research in any domain, the `cesR` package brings together these various sources, providing R users with a means of easily accessing each CES survey dataset while also circumventing the need to download and store survey data files. The structure of `cesR` also removes the need to call a data file from a directory, thereby making it easier to work with CES datasets between different workstations. Additionally, `cesR` provides a function, `get_decon()`, that creates a subset of the 2019 CES online survey that has been prepared in an opinionated way. This subset of the 2019 CES online survey provides a tool for educators that can be used to aid in the teaching of statistical analysis of large survey datasets.

The `cesR` package is built using the R programming language and is designed to work within the R integrated development environment (IDE) RStudio (R Core Team, 2020; RStudio Team, 2020). Providing access to data is common in R packages and the `cesR` package follows and was inspired by the work being done in the R community through such packages as the `opendatatoronto` package [@citeopendatatoronto], the `Lahman` package [@citeLahman], `fueleconomy` (Wickham, 2020), and `nasaweather` (Wickham, 2014). Furthermore, the `cesR` package is complementary to packages such as `dataverse` (Leeper, 2017) and `cancensus` (von Bergmann et al., 2020) that provide R users access to Canadian survey and census data. Packages such as these are important to R users as they improve the functionality of working within R by minimizing the number of steps required to load data and increasing the availability of data to R users. The `cesR` package contributes to this area of R package development and the utility of R by providing access to CES survey data while also providing a tool for educators.

@@

When called, the `get_ces()` function returns a requested CES survey as a data object and prints to the console the associated citation and URL for the survey dataset repository. The function takes one argument in the form of a character string. This argument is a vector member that has been associated with a CES survey through the body of code in the `get_ces()` function that when used calls the download URL for that survey on an associated GitHub repository named `ces_data`. If the provided character string argument matches a member of the built-in vector `ces_codes`, the associated file is downloaded using the `download.file()` function from the `utils` R package (R Core Team, 2020) as a compressed .zip folder and is stored temporarily in `inst/extdata` directory in the greater package directory. Upon downloading the file, the compressed folder is unzipped using the `unzip()` function from the `utils` R package (R Core Team, 2020) and read into R using either the `read_dta()` or `read_sav()` functions from the `haven` R package [@citehaven] depending on the file extension of the downloaded file. A data frame is then assigned using the `assign()` function  from the `base` R package (R Core Team, 2020) as a data object in the global environment. The downloaded file and file directory are then removed from the computer using the `unlink()` function from the `base` R package (R Core Team, 2020). Finally, the recommended citation for the requested survey dataset and URL of the survey data storage location are printed in the console (see [*2.1.1 `get_ces()` basics*](#211-get_ces-basics) for an example of a basic `get_ces()` function call).

@@

CES survey dataset files uses either a `.dta()` or `.sav()` file extension, meaning the datasets are loaded into R and RStudio as the type labelled. The `get_ces()` function does not convert the values of the loaded tables to a factor type so that personal workflow practices are not interfered with. It is suggested that to convert the dataset values to a factor type that the `to_factor()` function from the `labelled` package (Lamarange, 2020) be used (see [*2.1.3 `to_factor()` function*](#213-to_factor-function) for an example as to using the `to_factor()` function and `labelled` package).

@@

The `get_cescodes()` function works by constructing two vectors, one vector contains the CES survey codes and the other containing the associated survey argument calls. The function then creates dataframes of two vectors using the `data.frame()` function from the `base` package and adds an index number column using the `seq()` function from the `base` package (R Core Team, 2020) by which to merge the dataframes. Using the `merge()` function from the `base` package (R Core Team, 2020), the dataframes are then merged by the index number into a new dataframe. Using the index number ensures the dataframes remain ordered and merge correctly across rows. Column names in the new dataframe are then renamed using the `rename()` function from the `dplyr` package (Wickham et al., 2020) and the vector objects are removed from the environment. The `get_cescodes()` function does not create any variable that is available in the global environment.

## `get_question()`
The `get_question()` provides users with the ability to look up a survey question associated with a given column name. The function takes two arguments in the form of character strings, those being the name of a data object and the name of a column in the given data object. The function works such that it checks whether the given data object exists using the `exists()` function from the `base` package (R Core Team, 2020). If the object does not exist, the function will print out a warning in the console stating `Warning: Data object does not exist` (see [*2.3.2 `get_question()` data object error*](#232-get_question-data-object-error) for an example of this error). If the object does exist, `get_question()` will check if the given column name exists in the given data object. This is done using a combination of the `hasName()` function from the `utils` package and the `get()` function from the `base` package (R Core Team, 2020). The `hasName()` function checks if the given column name is in the given data object. Because the arguments are given as character strings the `get()` function is used to return the actual data object instead of the provided character string. Otherwise, the `hasName()` function would only check if the given column name argument occurred in the given character string argument and not the actual data object. If the column does not exist in the data object a warning is printed in the console stating `Warning: Variable is not in dataset`. See [*2.3.3 `get_question()` variable error*](#233-get_question-variable-error) for an example of this error message. If the given column exists in the given data object, `get_question()` will print the variable label of the given column to the console using a combination of the `var_label()` function from the `labelled` package (Larmarange, 2020) and the `get()` function from the `base` package (R Core Team, 2020). See [*2.3.1 `get_question()` basics*](#231-get_question-basics) for an example of the general use of the `get_question()` function and the [*3. Vignette*](#3-vignette) section for its use as part of the whole `cesR` package.

@@

In the initial argument classifications for the `get_preview()` function the variable `x` is set to equal six. This sets a default value for the number of rows to be returned if this value is left empty while allowing the value to be overwritten if provided a value. Like the [`get_ces()` function](#21-get_ces), the `get_preview()` function checks if a file exists in the cesR directory before downloading the requested survey data file associated with the given character string argument. If the file does not already exist, the survey data file is downloaded using the `download.file()` function from the `base` package to a temporary directory set using the `tempfile()` function also from the `base` package (R Core Team, 2020). The downloaded file is then unzipped into the `cesR` package directory using the `unzip()` function from the `utils` package (R Core Team, 2020). The extracted survey data file is read into R within the `get_preview()` function using either the `read_dta()` or `read_sav()` functions from the `haven` package depending upon the file extension (Wickham & Miller, 2020). The `get_preview()` function then creates a truncated version of the requested dataset by converting the dataset values to a factor type using the `to_factor()` function from the `labelled` package (Larmarange, 2020) and truncating the dataset using the `head()` function from the `utils` package (R Core Team, 2020) by the numeric value provided by the argument `x`. This dataset is assigned as a variable with the suffix `_preview` to the global environment using the `assign()` function from the `base` package (R Core Team, 2020). The function then removes the downloaded files using the `unlink()` function from the `base` package and removes the local variable holding the full dataset using the `rm()` function also from the `base` package (R Core Team, 2020). See [3. Vignette](#3-vignette) for a use of the `get_preview()` function in conjunction with the other `cesR` functions.

## get_decon()
When called, creates a subset of the 2019 CES online survey under the name `decon` (demographics and economics) that provides a tool for educators in the teaching of the analysis of large survey datasets. The `get_decon()` function takes no arguments. The function first checks the global environment if an object named `decon` exists using the `exists()` function from the `base` package (R Core Team, 2020). This prevents the `decon` dataset from being recreated if the object exists. If the `get_decon()` function is run when an object with the name `decon` already exists a warning will print in the console stating `Error in get_decon() : Warning: File already exists.` See [*2.5.1 `get_decon()` error*](#251-get_decon-error) for an example of the error message. If a situation arises in which the `decon` dataset needs to be recreated, then the best course of action is to use the `rm()` function from the `base` package (R Core Team, 2020) to remove the `decon` object and then run the `get_decon()` function again (see [*2.5.2 `get_decon()` reload*](#252-get_decon-reload)).

@@

After checking if an object with the name `decon` exists, the `get_decon()` function performs similarly to the `get_ces()` function. It assigns a temporary file extension with a .zip type using the `tempfile()` function from the `base` package, and calls on the URL for the 2019 CES online survey and temporarily downloads the associated file using the `download.file()` function from the `base` package (R Core Team, 2020). After unpacking the compressed folder with the `unzip()` function from the `utils` package (R Core Team, 2020), the file is read into R using the `read_dta()` function from the `haven` package (Wickham & Miller, 2020). Using the `select()` function from the `dplyr` package, 20 variables are selected from the main CES 2019 online survey and renamed using the `rename` function from the `dplyr` package (Wickham et al., 2020). The dataframe values are then converted to a factor type using the `to_factor()` function from the `labelled` package (Lamarange, 2020). A new variable, consisting of participants' responses to their self-perceived position on the political spectrum, is then created from two variables using the `unite()` function from the `tidyr` package (Wickham & Henry, 2020) and all empty cells are replaced with `NA` values using the `mutate()` function from the `dplyr` package (Wickham et al., 2020). The dataset is made available in the global environment by using the `assign()` function from the `base` package to create a new data object under the name `decon`. Lastly, the downloaded files are removed from the computer using the `unlink()` function from the `base` package (R Core Team, 2020) and a citation for the 2019 CES online survey and a URL to the storage location is printed to the console. See [*2.5.3 `get_decon()` basics*](#253-get_decon-basics) for an example of a general use of the `get_decon()` function.

@@

The following presents a vignette of installing `cesR` from its GitHub repository as well as calling and producing a subset of the 2019 CES phone survey dataset. This vignette uses functions from the `cesR`, `devtools` (Wickham et al., 2020), `labelled` (Larmarange, 2020), `dplyr` (Wickham et al., 2020), and `utils` (R Core Team, 2020) packages.

To begin, install and load the `cesR` package (and all other necessary packages) into RStudio. Currently, this is currently only available through the use of the `install_github` function from the `devtools` package (Wickham et al., 2020). During installation, RStudio may request to update other packages. It is best to press enter with an empty line (see [*3.1.1 Install `cesR`*](#311-install-cesR)).

@@

To check that the dataset has loaded into RStudio correctly, it is best to check the dataset with the `head()` function from the `utils` package (R Core Team, 2020). Unfortunately, the `head()` function does not work with labelled data, so it is best to convert the values to factors using the `to_factor()` function from the `labelled` package (Larmarange, 2020) (see [*3.1.5 Check 2019 CES phone data*](#315-check-ces-phone-data)).

@@

With the dataset loaded, the `select()` function from the `dplyr` package (Wickham et al., 2020) is an efficient way of selecting the required variable columns. In this case, the language, phone type, weight, citizenship, year of birth, interviewer gender, identified gender, province, satisfaction with Canadian democracy, and most important election issue are selected using the column index (see [*3.1.4 Select columns*](#314-select-columns)). Additionally, the columns are not all named in an understandable way and so will be renamed using the `rename()` function from the `dplyr` package (Wickham et al., 2020).

@@

We can see that some of the column names of the subset are not clearly named. The remedy this we can use the `rename` function from the `dplyr` package (Wickham et al., 2020). When using `rename()`, the order goes from new name to old name (see [*3.1.6 Rename columns*](#316-rename-columns)). To gain an idea of what to rename a column we can use the `get_question()` function to find the associated survey question (see [*3.1.5 `get_question()`*](#315-get_question)).

@@

# Next steps and cautions
The `cesR` package is dependent upon the functions from the `haven` (Wickham & Miller, 2020) and `labelled` (Larmarange, 2020) packages to be able to read the CES survey datasets. Thus, changes to the `haven` or `labelled` package may affect the functionality of the `cesR` package. As such, a future consideration would be to move away from this dependency to make the cesR package more independent and self-contained.

@@

# References

Canadian Election Study (2019). *CES Canadian Election Study EEC Etude electorale canadienne*. Canadian Election Study. http://www.ces-eec.ca/

Larmarange, J. (2020). labelled: Manipulating Labelled Data. R package version 2.5.0. https://CRAN.R-project.org/package=labelled

Leeper, T. J. (2017). dataverse: R Client for Dataverse 4. R package version 0.2.0.

Odesi (n.d.) *Scholars Portal*. http://odesi2.scholarsportal.info/webview/ _**(not sure if this citation/reference is correct. need to doublecheck)**_

R Core Team (2020). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

RStudio Team (2020). RStudio: Integrated Development Environment for R. RStudio, PBC, Boston, MA URL http://www.rstudio.com/.

UBC (n.d.). *Surveys*. Canadian Election Study Etude Electorale Canadienne. https://ces-eec.arts.ubc.ca/english-section/surveys/

von Bergmann, J., Dmitry Shkolnik, and Aaron Jacobs (2020). cancensus: R package to access, retrieve, and work with Canadian Census data and geography. v0.3.2.

Wickham, H. (2020). fueleconomy: EPA Fuel Economy Data. R package version 1.0.0. https://CRAN.R-project.org/package=fueleconomy

Wickham, H., François, R., Henry L., & Müller, K. (2020). dplyr: A Grammar of Data Manipulation. R package version 1.0.0. https://CRAN.R-project.org/package=dplyr

Wickham, H., & Henry, L. (2020). tidyr: Tidy Messy Data. R package version 1.1.0. https://CRAN.R-project.org/package=tidyr

Wickham, H., Hester, J., & Chang, W. (2020). devtools: Tools to Make Developing R Packages Easier. R package version 2.3.1. https://CRAN.R-project.org/package=devtools

Wickham, H., & Miller, E. (2020). haven: Import and Export 'SPSS', 'Stata' and 'SAS' Files. R package version 2.3.1. https://CRAN.R-project.org/package=haven

>>>>

An election season brings with it a stream of predictions made using a variety of resources and methods. While election predictions can provide insight as to the direction of an election and voter intention during an election, a retrospective on an election can be equally enlightening. Survey data of this sort can provide a better understanding of an election after-the-fact and can provide clarity as to the changes in the political climate over the years. The Canadian Election Study (CES) is a series of large-scale surveys conducted every election year that seeks to provide such a retrospective understanding to Canadian elections. By providing data on Canadian elections since 1965 up to, most recently, 2019 the CES provides insight into such topics as the intentions of voters, what issues voters deem important, and the perception of parties and candidates [@ces2019]. For example, @bittner2018leaders used CES data to assess the level of presidentialization experienced in Canadian elections over time; while other studies have used CES data to assess the level of Islamophobia in Canada [@wilkins2018islamophobia], how anti-party rhetoric affects voter behaviour [@gidengil2001correlates], and how political trust influences voter behaviour [@BelangerNadeau2005]. The collection of and access to such data is integral to this research. The R package `cesR` ('caesar'), as presented in this paper, seeks to complement the work of the CES, and enhance social science research and the teaching of statistical processes using survey data by providing a means for R users to easily access CES survey datasets.

The CES survey datasets are publicly available from various repositories across the internet in various data types ([@ces2019]; [@ubc]; [@odesi]). As the ability to find and access data is integral to quality research in any domain, the `cesR` package brings together these various sources, providing R users with a means of easily accessing each CES survey dataset while also circumventing the need to download and store survey data files. The structure of `cesR` also removes the need to call a data file from a directory, thereby making it easier to work with CES datasets between different workstations. Additionally, `cesR` provides a function, `get_decon()`, that creates a subset of the 2019 CES online survey that has been prepared in an opinionated way. This subset of the 2019 CES online survey provides a tool for educators that can be used to aid in the teaching of statistical analysis of large survey datasets.

The `cesR` package is built using the R programming language and is designed to work within the R integrated development environment (IDE) RStudio ([@citeR]; [@citeRStudio]). Providing access to data is common in R packages and the `cesR` package follows and was inspired by the work being done in the R community through such packages as the `opendatatoronto` package [@citeopendatatoronto], the `Lahman` package [@citeLahman], `fueleconomy` [@citefuel], and `nasaweather` [@citenasa]. Furthermore, the `cesR` package is complementary to packages such as `dataverse` [@citedataverse] and `cancensus` [@citecancensus] that provide R users access to Canadian survey and census data. Packages such as these are important to R users as they improve the functionality of working within R by minimizing the number of steps required to load data and increasing the availability of data to R users. The `cesR` package contributes to this area of R package development and the utility of R by providing access to CES survey data while also providing a tool for educators.

@@

When called, the `get_ces()` function returns a requested CES survey as a data object and prints to the console the associated citation and URL for the survey dataset repository. The function takes one argument in the form of a character string. This argument is a vector member that has been associated with a CES survey through the body of code in the `get_ces()` function that when used calls the download URL for that survey on an associated GitHub repository named `ces_data`. If the provided character string argument matches a member of the built-in vector `ces_codes`, the associated file is downloaded using the `download.file()` function from the `utils` R package [@citeR] as a compressed .zip folder and is stored temporarily in `inst/extdata` directory in the greater package directory. Upon downloading the file, the compressed folder is unzipped using the `unzip()` function from the `utils` R package [@citeR] and read into R using either the `read_dta()` or `read_sav()` functions from the `haven` R package [@citehaven] depending on the file extension of the downloaded file. A data frame is then assigned using the `assign()` function  from the `base` R package [@citeR] as a data object in the global environment. The downloaded file and file directory are then removed from the computer using the `unlink()` function from the `base` R package [@citeR]. Finally, the recommended citation for the requested survey dataset and URL of the survey data storage location are printed in the console (see [*2.1.1 `get_ces()` basics*](#211-get_ces-basics) for an example of a basic `get_ces()` function call).

@@

CES survey dataset files uses either a `.dta()` or `.sav()` file extension, meaning the datasets are loaded into R and RStudio as the type labelled. The `get_ces()` function does not convert the values of the loaded tables to a factor type so that personal work-flow practices are not interfered with. It is suggested that to convert the dataset values to a factor type that the `to_factor()` function from the `labelled` package [@citelabelled] be used (see [*2.1.3 `to_factor()` function*](#213-to_factor-function) for an example as to using the `to_factor()` function and `labelled` package).

@@

The `get_cescodes()` function works by constructing two vectors, one vector contains the CES survey codes and the other containing the associated survey argument calls. The function then creates dataframes of two vectors using the `data.frame()` function from the `base` package and adds an index number column using the `seq()` function from the `base` package [@citeR] by which to merge the dataframes. Using the `merge()` function from the `base` package [@citeR], the dataframes are then merged by the index number into a new dataframe. Using the index number ensures the dataframes remain ordered and merge correctly across rows. Column names in the new dataframe are then renamed using the `rename()` function from the `dplyr` package [@citedplyr] and the vector objects are removed from the environment. The `get_cescodes()` function does not create any variable that is available in the global environment.

## `get_question()`
The `get_question()` provides users with the ability to look up a survey question associated with a given column name. The function takes two arguments in the form of character strings, those being the name of a data object and the name of a column in the given data object. The function works such that it checks whether the given data object exists using the `exists()` function from the `base` package [@citeR]. If the object does not exist, the function will print out a warning in the console stating `Warning: Data object does not exist` (see [*2.3.2 `get_question()` data object error*](#232-get_question-data-object-error) for an example of this error). If the object does exist, `get_question()` will check if the given column name exists in the given data object. This is done using a combination of the `hasName()` function from the `utils` package and the `get()` function from the `base` package [@citeR]. The `hasName()` function checks if the given column name is in the given data object. Because the arguments are given as character strings the `get()` function is used to return the actual data object instead of the provided character string. Otherwise, the `hasName()` function would only check if the given column name argument occurred in the given character string argument and not the actual data object. If the column does not exist in the data object a warning is printed in the console stating `Warning: Variable is not in dataset`. See [*2.3.3 `get_question()` variable error*](#233-get_question-variable-error) for an example of this error message. If the given column exists in the given data object, `get_question()` will print the variable label of the given column to the console using a combination of the `var_label()` function from the `labelled` package [@citelabelled] and the `get()` function from the `base` package [@citeR]. See [*2.3.1 `get_question()` basics*](#231-get_question-basics) for an example of the general use of the `get_question()` function and the [*3. Vignette*](#3-vignette) section for its use as part of the whole `cesR` package.

@@

In the initial argument classifications for the `get_preview()` function the variable `x` is set to equal six. This sets a default value for the number of rows to be returned if this value is left empty while allowing the value to be overwritten if provided a value. Like the [`get_ces()` function](#21-get_ces), the `get_preview()` function checks if a file exists in the cesR directory before downloading the requested survey data file associated with the given character string argument. If the file does not already exist, the survey data file is downloaded using the `download.file()` function from the `base` package to a temporary directory set using the `tempfile()` function also from the `base` package [@citeR]. The downloaded file is then unzipped into the `cesR` package directory using the `unzip()` function from the `utils` package [@citeR]. The extracted survey data file is read into R within the `get_preview()` function using either the `read_dta()` or `read_sav()` functions from the `haven` package depending upon the file extension [@citehaven]. The `get_preview()` function then creates a truncated version of the requested dataset by converting the dataset values to a factor type using the `to_factor()` function from the `labelled` package [@citelabelled] and truncating the dataset using the `head()` function from the `utils` package [@citeR] by the numeric value provided by the argument `x`. This dataset is assigned as a variable with the suffix `_preview` to the global environment using the `assign()` function from the `base` package [@citeR]. The function then removes the downloaded files using the `unlink()` function from the `base` package and removes the local variable holding the full dataset using the `rm()` function also from the `base` package [@citeR]. See [3. Vignette](#3-vignette) for a use of the `get_preview()` function in conjunction with the other `cesR` functions.

## get_decon()
When called, creates a subset of the 2019 CES online survey under the name `decon` (demographics and economics) that provides a tool for educators in the teaching of the analysis of large survey datasets. The `get_decon()` function takes no arguments. The function first checks the global environment if an object named `decon` exists using the `exists()` function from the `base` package [@citeR]. This prevents the `decon` dataset from being recreated if the object exists. If the `get_decon()` function is run when an object with the name `decon` already exists a warning will print in the console stating `Error in get_decon() : Warning: File already exists.` See [*2.5.1 `get_decon()` error*](#251-get_decon-error) for an example of the error message. If a situation arises in which the `decon` dataset needs to be recreated, then the best course of action is to use the `rm()` function from the `base` package [@citeR] to remove the `decon` object and then run the `get_decon()` function again (see [*2.5.2 `get_decon()` reload*](#252-get_decon-reload)).

@@

After checking if an object with the name `decon` exists, the `get_decon()` function performs similarly to the `get_ces()` function. It assigns a temporary file extension with a .zip type using the `tempfile()` function from the `base` package, and calls on the URL for the 2019 CES online survey and temporarily downloads the associated file using the `download.file()` function from the `base` package [@citeR]. After unpacking the compressed folder with the `unzip()` function from the `utils` package [@citeR], the file is read into R using the `read_dta()` function from the `haven` package [@citehaven]. Using the `select()` function from the `dplyr` package, 20 variables are selected from the main CES 2019 online survey and renamed using the `rename` function from the `dplyr` package [@citedplyr]. The dataframe values are then converted to a factor type using the `to_factor()` function from the `labelled` package [@citelabelled]. A new variable, consisting of participants' responses to their self-perceived position on the political spectrum, is then created from two variables using the `unite()` function from the `tidyr` package [@citetidyr] and all empty cells are replaced with `NA` values using the `mutate()` function from the `dplyr` package [@citedplyr]. The dataset is made available in the global environment by using the `assign()` function from the `base` package to create a new data object under the name `decon`. Lastly, the downloaded files are removed from the computer using the `unlink()` function from the `base` package [@citeR] and a citation for the 2019 CES online survey and a URL to the storage location is printed to the console. See [*2.5.3 `get_decon()` basics*](#253-get_decon-basics) for an example of a general use of the `get_decon()` function.

@@

The following presents a vignette of installing `cesR` from its GitHub repository as well as calling and producing a subset of the 2019 CES phone survey dataset. This vignette uses functions from the `cesR`, `devtools` [@citedevtools], `labelled` [@citelabelled], `dplyr` [@citedplyr], and `utils` [@citeR] packages.

To begin, install and load the `cesR` package (and all other necessary packages) into RStudio. Currently, this is currently only available through the use of the `install_github` function from the `devtools` package [@citedevtools]. During installation, RStudio may request to update other packages. It is best to press enter with an empty line (see [*3.1.1 Install `cesR`*](#311-install-cesR)).

@@

To check that the dataset has loaded into RStudio correctly, it is best to check the dataset with the `head()` function from the `utils` package [@citeR]. Unfortunately, the `head()` function does not work with labelled data, so it is best to convert the values to factors using the `to_factor()` function from the `labelled` package [@citelabelled] (see [*3.1.5 Check 2019 CES phone data*](#315-check-ces-phone-data)).

@@

With the dataset loaded, the `select()` function from the `dplyr` package [@citedplyr] is an efficient way of selecting the required variable columns. In this case, the language, phone type, weight, citizenship, year of birth, interviewer gender, identified gender, province, satisfaction with Canadian democracy, and most important election issue are selected using the column index (see [*3.1.4 Select columns*](#314-select-columns)). Additionally, the columns are not all named in an understandable way and so will be renamed using the `rename()` function from the `dplyr` package [@citedplyr].

@@

We can see that some of the column names of the subset are not clearly named. The remedy this we can use the `rename` function from the `dplyr` package [@citedplyr]. When using `rename()`, the order goes from new name to old name (see [*3.1.6 Rename columns*](#316-rename-columns)). To gain an idea of what to rename a column we can use the `get_question()` function to find the associated survey question (see [*3.1.5 `get_question()`*](#315-get_question)).

@@

# Next steps and cautions
The `cesR` package is dependent upon the functions from the `haven` [@citehaven] and `labelled` [@citelabelled] packages to be able to read the CES survey datasets. Thus, changes to the `haven` or `labelled` package may affect the functionality of the `cesR` package. As such, a future consideration would be to move away from this dependency to make the cesR package more independent and self-contained.

@@
